---
title: "Ingalls Standards Project"
author: "R.Lionheart"
date: "Start date: 3/17/2020"
output:
  html_document: default
  pdf_document: default
---

### Overview
The original version of the Ingalls_Lab_Standards.csv lives in Katherine's GitHub at the following link: 
[Ingalls Lab Standards](https://github.com/kheal/Example_Untargeted_Metabolomics_Workflow)

A new repository has been made for this project: [New Ingalls Lab Standards](https://github.com/IngallsLabUW/Ingalls_Standards). This repository contains all the modifications and source code the Ingalls Standards csv, as well as all previous iterations of the csv.

View the most recent Ingalls Standards in this repository, titled Ingalls_Standards_NEW.csv.

------

#### Changes made in v1.0

*Capitalization*  
* All "Acid"s changed to "acid", as per KEGG conventions. Example: 3-Sulfopyruvic Acid --> 3-Sulfopyruvic acid   
* First appearance of compound in complete name capitalized. Example: 7-dehydrocholesterol --> 7-Dehydrocholesterol, thiamine pyrophosphate --> Thiamine pyrophosphate  
* "B-compoundName" changed to beta-CompoundName. Example: B-ionine --> beta_Ionine  
* Second part of compound name lowercase. Example: Glutathione Disulfide --> Glutathione disulfide  

*Abbreviations*  
To alleviate long commpound names, the primary ID will remain as the KEGG name or reasonable abbreviation but a more manageable name will go under the "Figure Names" column. 

*Symbols*  
All symbols that are part of the KEGG name will remain in place. A syntactically correct column will be added for upload to CMAP, etc.

*Other specific changes*  
* RP B12 --> Cyanocobalamin  
* All B12s --> Methylcobalamin, Hydroxocobalamin, etc.  
* Betaine --> Glycine betaine  
* Dimethyl glycine --> Dimethylglycine  
* All Vitamins changed to their descriptive name: Vitamine B1 --> Thiamine, Vitamin B2 --> Riboflavin, etc.  


#### Changes made in v2.0

*Figure Names column added*
* Add in a Figure Name column.  
* Replace vitamins with full names.
* Bug fixes in primary and figure name columns.


#### Changes made in v3.0


#### Long-term Changes
* Include a column for syntactically correct names, which may or may not be the primary name.
* Possibly include a column of annotation, depending on size/how much info is in that column.
* Set up system to keep note of changes to official KEGG compound names whether or not those changes are enacted.  
* Add stereoisomer orientation column.  
* List of all commonly used names for a compound (eg Glutamic acid == Glutamate).  

------

## Change Log
4/8/20: Internal standards updated to reflect consistent naming scheme.  
4/20/20: Abbreviated compounds switched to full-length KEGG names.  
4/20/20: Adjusted compounds with incorrect capitalization.  
4/22/20: Added or removed symbols as required by KEGG names.  
4/23/20: Update last compounds not falling into the above groups.
4/23/20: Complete renaming of compounds, including those that didn't change from the original names.
4/28/20: Adjust vitmins to descriptive names.
4/29/20: Adjust abbreviations for consistency.
5/8/20: FINAL NAMES DECIDED FOR PRIMARY ID.
5/28/20: Bug fixes in final names.
5/28/20: First suggestions for Figure Name column.
6/3/20: Final Figure Names decided.
7/20/20: ChEBI IDs modified, new column added.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("src/Functions.R")

library(stringr)
library(tidyverse)
```

------

```{r Source Primary ID code, include=TRUE}
matching.variable = "Ingalls"
filenames <- RemoveCsv(list.files(path = "data_raw", pattern = matching.variable))

for (i in filenames) {
  filepath <- file.path("data_raw", paste(i, ".csv", sep = ""))
  assign(make.names(i), read.csv(filepath, stringsAsFactors = FALSE, fileEncoding = "latin1"))
}
source("src/Primary_ID.R")

write.csv(Ingalls_Lab_Standards_Extras, "Ingalls_Lab_Standards_NEW.csv")

```

-----

```{r Source Figures Names code, incldue=TRUE}

source("src/Figure_Names.R")

write.csv(Ingalls_Lab_Standards_FigNames, "Ingalls_Lab_Standards_NEW.csv")

```

