---
title: "Ingalls Standards Project"
author: "R.Lionheart"
date: "Start date: 3/17/2020"
output:
  html_document: default
  pdf_document: default
---

### Overview
The most updated version of the Ingalls_Lab_Standards.csv lives in Katherine's github at the following link: 
[Ingalls Lab Standards](https://github.com/kheal/Example_Untargeted_Metabolomics_Workflow)

A new repository has been made for this project: [New Ingalls Lab Standards](https://github.com/IngallsLabUW/Ingalls_Standards)

View the most recent Ingalls Standards in this repository, titled Ingalls_Standards_NEW.csv.

------

#### Changes made in v1.0

*Capitalization*  
* All "Acid"s changed to "acid", as per KEGG conventions. Example: 3-Sulfopyruvic Acid --> 3-Sulfopyruvic acid   
* First appearance of compound in complete name capitalized. Example: 7-dehydrocholesterol --> 7-Dehydrocholesterol, thiamine pyrophosphate --> Thiamine pyrophosphate  
* "B-compoundName" changed to beta-CompoundName. Example: B-ionine --> beta_Ionine  
* Second part of compound name lowercase. Example: Glutathione Disulfide --> Glutathione disulfide  

*Abbreviations*  
To alleviate long commpound names, the primary ID will remain as the KEGG name or reasonable abbreviation but a more manageable name will go under the "Figure Names" column. 

*Symbols*  
All symbols that are part of the KEGG name will remain in place. A syntactically correct column will be added for upload to CMAP, etc.

*Other specific changes*  
* RP B12 --> Cyanocobalamin  
* All B12s --> Methylcobalamin, Hydroxocobalamin, etc.  
* Betaine --> Glycine betaine  
* Dimethyl glycine --> Dimethylglycine  
* All Vitamins changed to their descriptive name: Vitamine B1 --> Thiamine, Vitamin B2 --> Riboflavin, etc.  


#### Long-term Changes
* Add in a Figure Name column.  
* Include a column for syntactically correct names, which may or may not be the primary name.
* Possibly include a column of annotation, depending on size/how much info is in that column.
* Set up system to keep note of changes to official KEGG compound names whether or not those changes are enacted.  
* Add stereoisomer orientation column.  
* List of all commonly used names for a compound (eg Glutamic acid == Glutamate).  

------

## Change Log
4/8/20: Internal standards updated to reflect consistent naming scheme.  
4/20/20: Abbreviated compounds switched to full-length KEGG names.  
4/20/20: Adjusted compounds with incorrect capitalization.  
4/22/20: Added or removed symbols as required by KEGG names.  
4/23/20: Update last compounds not falling into the above groups.
4/23/20: Complete renaming of compounds, including those that didn't change from the original names.
4/28/20: Adjust vitmins to descriptive names.
4/29/20: Adjust abbreviations for consistency.
5/8/20: FINAL NAMES DECIDED FOR PRIMARY ID.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("src/Functions.R")

library(stringr)
library(tidyverse)
```

------

```{r Import Internal Standards, include=FALSE}
matching.variable = "Ingalls"
filenames <- RemoveCsv(list.files(path = "data_raw", pattern = matching.variable))

for (i in filenames) {
  filepath <- file.path("data_raw", paste(i, ".csv", sep = ""))
  assign(make.names(i), read.csv(filepath, stringsAsFactors = FALSE, fileEncoding = "latin1"))
}
```

## Import edits from LTC.
### Renaming of typos/specific issues.
```{r Change Internal Standards, include=TRUE}
# View list of all changed Internal Standards.
LauraEditsIS <- Ingalls_Lab_Standards_LauraEdit %>%
  select(Compound.Type,Compound.Name_new, Compound.Name_old) %>%
  filter(Compound.Type == "Internal Standard") %>%
  filter(!Compound.Name_new == Compound.Name_old) # Remove compounds that don't change

Ingalls_Lab_Standards_LauraEdit[Ingalls_Lab_Standards_LauraEdit == "O-Propanoylcarnitine"] <- "O-Propionylcarnitine"
Ingalls_Lab_Standards_LauraEdit[Ingalls_Lab_Standards_LauraEdit == "O-Acetyl-L-carnitine"] <- "O-Acetylcarnitine"

knitr::kable(LauraEditsIS[1:10, ], caption="Changed Internal Standards", floating.environment="sidewaystable")
```

## Renaming of Internal Standards
```{r Incorporate IS changes, include=TRUE}
Ingalls_Lab_Standards_IS <- Ingalls_Lab_Standards %>%
  left_join(Ingalls_Lab_Standards_LauraEdit %>% rename(Compound.Name = Compound.Name_old)) %>%
  rename(Compound.Name_old = Compound.Name,
         Compound.Name = Compound.Name_new) %>%
  select(Compound.Type, Column, Compound.Name, Compound.Name_old, everything()) %>%
  mutate(Compound.Name_old = 
           ifelse(Compound.Type == "Internal Standard" & Compound.Name == Compound.Name_old, 
                                    Compound.Name, Compound.Name_old)) %>%
  mutate(Compound.Name = ifelse(Compound.Type == "Internal Standard", Compound.Name, NA))

write.csv(Ingalls_Lab_Standards_IS, "Ingalls_Lab_Standards_NEW.csv")
```

-----

## Abbreviations
```{r Next experimental Laura, include=FALSE}
Abbreviations <- Ingalls_Lab_Standards_LauraEdit %>%
  select(Compound.Name_new, Compound.Name_old, Compound.Type) %>% 
  mutate(Compound.Name_new = recode(Compound.Name_new,
                                    "DMSP" = "Dimethylsulfoniopropionate",
                                    "ADP" = "Adenosine diphosphate",
                                    "AMP" = "Adenosine monophosphate",
                                    "ATP" = "Adenosine triphosphate",
                                    "FAD" = "Flavin adenine dinucleotide",
                                    "GMP" = "Guanosine monophosphate",
                                    "GMP, 15N5" = "Guanosine monophosphate, 15N5",
                                    "GTP" = "Guanosine triphosphate",
                                    "PEP" = "Phosphoenolpyruvic acid",
                                    "TMAB" = "(3-Carboxypropyl)trimethylammonium",
                                    "cAMP" = "3',5'-Cyclic AMP",
                                    "cGMP" = "3',5'-Cyclic GMP",
                                    "Cys-Gly" = "L-Cysteinylglycine")) %>%
  filter(Compound.Type != "Internal Standard",
         Compound.Name_new != Compound.Name_old) %>% 
  mutate(letter.count = nchar(Compound.Name_old)) %>%
  arrange(letter.count) %>%
  filter(nchar(Compound.Name_old) < 10 | Compound.Name_old == "(3-Carboxypropyl)trimethylammonium (TMAB)",
         nchar(Compound.Name_old) < 7 | str_detect(Compound.Name_old, "-"))
```

### Show list of abbreviated compounds to be changed
```{r View abbreviation table, include=TRUE}
knitr::kable(Abbreviations, caption="Potential Abbreviation Changes", floating.environment="sidewaystable")
```

### Finalize changes of abbreviated compounds
```{r Change abbreviated compounds, include=TRUE}
Ingalls_Lab_Standards_Abbr <- Ingalls_Lab_Standards_IS %>%
  left_join(Abbreviations %>% select(Compound.Name_new, Compound.Name_old) %>% unique()) %>%
  select(Compound.Name, Compound.Name_new, Compound.Name_old, Compound.Type, everything()) %>%
  mutate(Compound.Name = ifelse(is.na(Compound.Name), Compound.Name_new, Compound.Name)) %>%
  select(Compound.Type, Column, everything(), -Compound.Name_new) 

write.csv(Ingalls_Lab_Standards_IS, "Ingalls_Lab_Standards_NEW.csv")
```

-----

### Isolate different capitalizations of compounds
```{r Capitalizations, include=FALSE}
Capitalizations <- Ingalls_Lab_Standards_LauraEdit %>%
  select(Compound.Name_new, Compound.Name_old) %>%
  filter(!Compound.Name_new %in% Ingalls_Lab_Standards_Abbr$Compound.Name, 
          Compound.Name_new != Compound.Name_old) # drop commpounds already reassigned or matching originals
```


```{r View Capitalizations, include=TRUE}
Acid.Capitals <- Capitalizations %>%
  filter(str_detect(Compound.Name_new, regex("acid", ignore_case = TRUE))) %>%
  select(Compound.Name_new, Compound.Name_old)
Beta.Capitals <- Capitalizations %>%
  filter(str_detect(Compound.Name_new, regex("beta-", ignore_case = TRUE))) %>%
  select(Compound.Name_new, Compound.Name_old)
Just.Capitals <- Capitalizations %>%
  select(Compound.Name_new, Compound.Name_old) %>%
  filter(!Compound.Name_new %in% c(Acid.Capitals$Compound.Name_new, Beta.Capitals$Compound.Name_new),
         tolower(Compound.Name_new) == tolower(Compound.Name_old),
         Compound.Name_new != Compound.Name_old) %>%
  unique()

```

### View "Acids", "Beta", and different-capitalization compounds.
```{r View capitalization tables, include=TRUE}
Complete.Caps <- Acid.Capitals %>%
  rbind(Beta.Capitals) %>%
  rbind(Just.Capitals)
```

### Finalize changes of capitalized compounds
```{r Change capitalized compounds, include=TRUE}
Ingalls_Lab_Standards_Caps <- Ingalls_Lab_Standards_Abbr %>%
  left_join(Complete.Caps) %>%
  select(Compound.Name, Compound.Name_new, Compound.Name_old, Compound.Type, everything()) %>%
  mutate(Compound.Name = ifelse(is.na(Compound.Name), Compound.Name_new, Compound.Name)) %>%
  select(Compound.Type, Column, everything(), -Compound.Name_new) %>%
  unique()

write.csv(Ingalls_Lab_Standards_Caps, "Ingalls_Lab_Standards_NEW.csv")
```

-----

```{r Symbols, include=FALSE}
Symbols <- Ingalls_Lab_Standards_Caps %>%
  left_join(Ingalls_Lab_Standards_LauraEdit) %>%
  select(Compound.Name, Compound.Name_new, Compound.Name_old, Compound.Type) %>%
  filter(is.na(Compound.Name)) %>%
  filter(Compound.Name_new != Compound.Name_old) %>%
  select(Compound.Name_new, Compound.Name_old) %>%
  filter_all(any_vars(str_detect(., "[^[:alnum:] ]"))) %>%
  unique()

```

### View Symbols that will be changed
```{r View symbols table, include=TRUE}
knitr::kable(Symbols[1:10, ], caption="First 10 Symbol Changes", floating.environment="sidewaystable")
```

### Finalize changes of compounds with symbols
```{r Change compounds with symbols, include=TRUE}
Ingalls_Lab_Standards_Symbols <- Ingalls_Lab_Standards_Caps %>%
  left_join(Symbols) %>%
  select(Compound.Name, Compound.Name_new, Compound.Name_old, Compound.Type, everything()) %>%
  mutate(Compound.Name = ifelse(is.na(Compound.Name), Compound.Name_new, Compound.Name)) %>%
  select(Compound.Type, Column, everything(), -Compound.Name_new) %>%
  unique()

write.csv(Ingalls_Lab_Standards_Symbols, "Ingalls_Lab_Standards_NEW.csv")
```

------

### Adjust Vitamins to descriptive names
```{r Vitamins, include=FALSE}
#Ingalls_Lab_Standards_Vitamins <- Ingalls_Lab_Standards_Extras %>% ORIGINAL
Ingalls_Lab_Standards_Vitamins <- Ingalls_Lab_Standards_Symbols %>%
  mutate(Compound.Name = recode(Compound.Name, 
          "Vitamin B2, 13C4, 15N2" = "Riboflavin-dioxopyrimidine, 13C4, 15N2",
         	"Vitamin B2" = "Riboflavin",
          "Vitamin C" = "Ascorbic acid",
          "Vitamin D2" = "Calciferol",
          "Vitamin K1" = "Phytonadione",
          "Vitamin K2" = "Menaquinone",
          "AMP, 15N5" = "Adenosine monophosphate, 15N5",
          "Methyl (indol-3-yl)acetate" = "Indole-3-methyl acetate")) %>%
  mutate(Compound.Name = ifelse(Compound.Name_old == "Methyl indole 3 carboxylate", "Methyl indole-3-carboxylate", Compound.Name))

write.csv(Ingalls_Lab_Standards_Vitamins, "Ingalls_Lab_Standards_NEW.csv")

```


### All other compounds not yet changed
```{r Currently unchanged compounds, include=FALSE}
Unchanged <- Ingalls_Lab_Standards_Vitamins %>%
  left_join(Ingalls_Lab_Standards_LauraEdit) %>%
  select(Compound.Name, Compound.Name_new, Compound.Name_old) %>%
  filter(is.na(Compound.Name)) %>%
  filter(Compound.Name_new != Compound.Name_old)

```


### Add changes of still-unedited compounds - Including those that remain as the original name.
```{r Add unchanged compounds, include=TRUE}
Ingalls_Lab_Standards_Extras <- Ingalls_Lab_Standards_Vitamins %>%
  left_join(Unchanged) %>%
  select(Compound.Name, Compound.Name_new, Compound.Name_old, Compound.Type, everything()) %>%
  mutate(Compound.Name = ifelse(is.na(Compound.Name), Compound.Name_new, Compound.Name)) %>%
  mutate(Compound.Name = ifelse(is.na(Compound.Name), Compound.Name_old, Compound.Name)) %>%
  select(Compound.Type, Column, everything(), -Compound.Name_new) %>%
  unique()

write.csv(Ingalls_Lab_Standards_Extras, "Ingalls_Lab_Standards_NEW.csv")
```
