---
title: "Ingalls Standards Project"
author: "R.Lionheart"
date: "Start date: 3/17/2020"
output:
  html_document: default
  pdf_document: default
---

### Overview
The most updated version of the Ingalls_Lab_Standards.csv lives in Katherine's github at the following link: 
[Ingalls Lab Standards](https://github.com/kheal/Example_Untargeted_Metabolomics_Workflow)

A new repository has been made for this project: [New Ingalls Lab Standards](https://github.com/IngallsLabUW/Ingalls_Standards)


View the most recent Ingalls Standards in this repository, titled Ingalls_Standards_NEW.csv.

------

#### Proposed short-term changes

#### Internal Standard Changes: 
1. Standardize the notation for Internal Standards. Potential changes completed by LTC on 04/02/20.
    
    
2. Enact consistent capitalization. 
    + Use regex syntax to ignore case when matching to real data.
    + Check on intra-comound capitalization, eg "1-stearoyl-2-oleoyl-sn-glycero-3-phosphoethanolamine"
    + Make sure compound capitalization doesn't interfere with the KEGG primary name.
3. Remove symbols, such as question marks, dashes when they're not needed. 
    + Keep/include any symbols that are part of the KEGG name.


**Examples of Specific Compounds for first changes**
This will include the actual outcome for each compound that should be changed. It may be part of a larger sheet that includes a column of "Before" and "After" compound changes so the new names can be checked against legacy data. 

*Symbols*  
* B-apo-8'-carotenal  
* Keto-?-(methylthio)butyric acid  
* N(e)-Acetyl-Lysine  
* Trimethylammonium Propionate (TMAP)  

*Capitalization*  
See below for full details  
* Dimethyl glycine, Dimethyl Glycine  
* thiamine pyrophosphate, Thiamine pyrophosphate  

*Others*  
* RP B12, which exists in addition to plain B12. This should be Cyano b12 or Cyanocobalamin. This is a leftover mistake from replacing cyano to RP for the chromatography.                                     

------

#### Long-term Changes
* Primary name column, how we refer to it in the lab and how it will be ID'd in analysis.
* Include a column for syntactically correct names, which may or may not be the primary name.
* Possibly include a column of annotation, depending on size/how much info is in that column.

------

# Change Log
4/8/20: Internal standards updated to reflect consistent naming scheme.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("src/Functions.R")
library(tidyverse)
```

```{r imports, include=FALSE}
Kegg.compounds <- read.csv("data_raw/KEGGCompounds_withMasses.csv", stringsAsFactors = F)
Ingalls.compounds <- read.csv("data_raw/Ingalls_Lab_Standards.csv", stringsAsFactors = F)

kegg.names <- Kegg.compounds %>%
  arrange(Kegg.compounds$Name) %>%
  select(Name, AllNames)
ingalls.names <- sort(unique(Ingalls.compounds$Compound.Name))
```

## Renaming of Internal Standards
```{r Import Internal Standards, include=FALSE}
# Assign matching variable to 
matching.variable = "Ingalls"
filenames <- RemoveCsv(list.files(path = "data_raw", pattern = matching.variable))

for (i in filenames) {
  filepath <- file.path("data_raw", paste(i, ".csv", sep = ""))
  assign(make.names(i), read.csv(filepath, stringsAsFactors = FALSE, fileEncoding = "latin1"))
}
```

```{r Rename Internal Standards, include=TRUE}
# View list of all changed Internal Standards.
LauraEditsIS <- Ingalls_Lab_Standards_LauraEdit %>%
  select(Compound.Type,Compound.Name_new, Compound.Name_old) %>%
  filter(Compound.Type == "Internal Standard") %>%
  filter(!Compound.Name_new == Compound.Name_old) # Remove compounds that don't change

knitr::kable(LauraEditsIS[c(2:3)],
            caption="Renamed Ingalls Standards", floating.environment="sidewaystable")
```

```{r Incorporate IS changes, include=TRUE}
Ingalls_Lab_Standards_IS <- Ingalls_Lab_Standards %>%
  left_join(LauraEditsIS %>% rename(Compound.Name = Compound.Name_old)) %>%
  mutate(Compound.Name = ifelse(is.na(Compound.Name_new), Compound.Name, Compound.Name_new))

write.csv(Ingalls_Lab_Standards_IS, "Ingalls_Lab_Standards_NEW.csv")

```

<!-- ##### Compounds that are an exact match between KEGG compounds and Ingalls Standards compounds. -->
<!-- ```{r matched names, include = TRUE} -->
<!-- matched <- Kegg.compounds %>% filter(Name %in% ingalls.names) -->
<!-- knitr::kable(matched[, 4:6],caption="Current KEGG/Ingalls matched compounds", floating.environment="sidewaystable") -->
<!-- ``` -->

<!-- ##### Compounds that are duplicated within the Ingalls_Lab_Standards csv, for HILIC ion mode or multiple run type detection. -->
<!-- ```{r duplicates, include=TRUE} -->
<!-- duplicated.ingalls <- Ingalls.compounds %>% -->
<!--   group_by(Category = stringi::stri_trans_totitle(Compound.Name)) %>% -->
<!--   filter(n() > 1) %>% -->
<!--   arrange(Compound.Name) -->

<!-- knitr::kable(duplicated.ingalls[c(2:3, 10, 16:17, 19, 22)],  -->
<!--              caption="Duplicated Ingalls Compounds", floating.environment="sidewaystable") -->
<!-- ``` -->

<!-- ------ -->

<!-- ### Code + Analysis begins here -->

<!-- #### All Ingalls standards capitalized.  -->
<!-- This doesn't account for those compounds that don't need to be capitalized, such as p-Coumaric acid, which now shows up as P-Coumaric Acid. This will be edited to reflect the KEGG compound primary name. -->
<!-- ```{r experimental, include=FALSE} -->
<!-- CapitalizeWords <- function(x) { -->
<!--   s <- strsplit(x, " ")[[1]] -->
<!--    paste(toupper(substring(s, 1, 1)), substring(s, 2), -->
<!--        sep="", collapse=" ") -->
<!-- } -->

<!-- cmpds.capitalized <- as.data.frame(sapply(ingalls.names, CapitalizeWords)) %>% rownames_to_column() -->
<!-- colnames(cmpds.capitalized) <- c("Original", "Capitalized") -->
<!-- cmpds.changed <- cmpds.capitalized[, 1:2] %>% -->
<!--   mutate(Change = rowSums(cmpds.capitalized == cmpds.capitalized[, 1]) == ncol(cmpds.capitalized)) %>% -->
<!--   filter(Change == FALSE) %>% -->
<!--   select(-Change) -->
<!-- ``` -->

<!-- ```{r capitalize, include=TRUE} -->
<!-- knitr::kable(cmpds.changed,  -->
<!--              caption="All Compounds Capitalized", floating.environment="sidewaystable") -->
<!-- ``` -->

<!-- Check for improper changes against the KEGG compounds. This returns row locations of all matches. Will require some work to filter compounds correctly. -->
<!-- ```{r KEGG check, include=TRUE} -->

<!-- ethanol.example <- sapply(colnames(kegg.names[]),  -->
<!--                           function(x) grep("ethanol", kegg.names[, x], ignore.case = TRUE)) -->

<!-- ethanol.example -->

<!-- ``` -->
