---
title: "Ingalls Standards Project: Structural Changes"
author: "R.Lionheart"
date: "Start date: 3/17/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("src/Functions.R")

library(stringr)
library(tidyverse)
```

This markdown document controls the original large structural changes initiated at the beginning of the repository (March 2020). "Large Structural Changes" include the creation of new columns, multi-group name changes of compounds, etc.

## Change Log
4/8/20: Internal standards updated to reflect consistent naming scheme.  
4/20/20: Abbreviated compounds switched to full-length KEGG names.  
4/20/20: Adjusted compounds with incorrect capitalization.  
4/22/20: Added or removed symbols as required by KEGG names.  
4/23/20: Update last compounds not falling into the above groups.
4/23/20: Complete renaming of compounds, including those that didn't change from the original names.
4/28/20: Adjust vitamins to descriptive names.
4/29/20: Adjust abbreviations for consistency.
5/8/20: FINAL NAMES DECIDED FOR PRIMARY ID.
5/28/20: Bug fixes in final names.
5/28/20: First suggestions for Figure Name column.
6/3/20: Final Figure Names decided.
7/30/20: ChEBI names scraped from website and added in new column.
8/3/20: IUPAC, InCHI, and SMILES columns added to standards.
8/10/20: Classyfire classifications added for those compounds that have them.
10/14/20: First version of SQL-safe names introduced. 
10/22/20: Remove all trailing numbers from SQL-safe names. This is the column to be used for CMAP.
------

```{r Source Primary ID code, include=TRUE}
matching.variable = "Ingalls"
filenames <- RemoveCsv(list.files(path = "data_raw", pattern = matching.variable))

for (i in filenames) {
  filepath <- file.path("data_raw", paste(i, ".csv", sep = ""))
  assign(make.names(i), read.csv(filepath, stringsAsFactors = FALSE, fileEncoding = "latin1"))
}

source("src/Primary_ID.R")

write.csv(Ingalls_Lab_Standards_Extras, "Ingalls_Lab_Standards_NEW.csv")

```

```{r Source Figures Names code, include=TRUE}

source("src/Figure_Names.R")

write.csv(Ingalls_Lab_Standards_FigNames, "Ingalls_Lab_Standards_NEW.csv")
```

-----

```{r Source ChEBI code, include=TRUE}

source("src/CHEBI_check.R")

write.csv(Ingalls_Lab_Standards_ChEBI, "Ingalls_Lab_Standards_NEW.csv")
```

-----

```{r Source IUPAC, InCHI, and SMILES code, include=TRUE}

source("src/IUPAC_InChl_Smiles.R")

write.csv(Ingalls_Lab_Standards_IUPAC.InCHI.SMILES, "Ingalls_Lab_Standards_NEW.csv")
```

-----

The next section deals with the ClassyFire ID section. 
The source code for creating the ClassyFire IDs commented out: *DO NOT RUN THE High Class Function LINE UNLESS YOU ARE STARTING THE PROJECT FROM SCRATCH!!* It can take several hours just to complete.

```{r Source Classyfire code, include=TRUE}

### source("src/High_Class_Function_062220.R") ###
source("src/Classyfire_column.R")

Ingalls_Lab_Standards_Classyfire$m.z <- sub("188.055899Â", "188.055899",
                                            Ingalls_Lab_Standards_Classyfire$m.z)

write.csv(Ingalls_Lab_Standards_Classyfire, "Ingalls_Lab_Standards_NEW.csv", row.names = FALSE)
```

-----

```{r Source SQL-Safe code, include=TRUE}

source("src/SQL_Safe.R")
Ingalls_Lab_Standards_SQLSafe$Compound.Name_SQL <- gsub("Vit", "Vitamin", Ingalls_Lab_Standards_SQLSafe$Compound.Name_SQL)
Ingalls_Lab_Standards_SQLSafe <- Ingalls_Lab_Standards_SQLSafe %>%
  mutate(Compound.Name_SQL = recode(Compound.Name_SQL,
                                    "Vitaminamin B2 13C4 15N2" = "Vitamin B12 13C4 15N2"))

write.csv(Ingalls_Lab_Standards_SQLSafe, "Ingalls_Lab_Standards_NEW.csv", row.names = FALSE)
```

